---
title: "Predict the excercise activity using the data collected by the wearable devices"
output:
  md_document
---
### Special thanks to web archive http://web.archive.org/ for allwoing to use the datasets
### This report shows the predictive analysis using Machine learning algorithm for the 20 cases in the  excercise data 
### Data is collected  from accelerometers on the belt, forearm, arm, and dumbell of 6 participants
### Load the training and test data 
* Note: Reduced data set used as the PC where this report ran is having less computing power *

```{r load, include=TRUE}
fpath <- "F:/c8proj/"
  trainfname <- "pml-training.csv"
  setwd(fpath)
  testfname <- "pml-testing.csv"
  
  if (!file.exists(trainfname)){
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv",trainfname)
  }
  
  if (!file.exists(testfname)){
  download.file("https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv",testfname)
    }
  
  traindset <- read.csv("pml-training.csv",header = TRUE , nrows = 6000 ,na.strings = c("NA", ""))
  testdset <- read.csv("pml-testing.csv",header = TRUE,na.strings = c("NA", ""))
```
### Load all required libraries Caret, rpart , random forest
```{r loadlib, include=TRUE}
library(caret)
library(rpart)
library(randomForest)
library(repmis)
```
### Clean and split the data into test and validation partitions
```{r Split, include=TRUE}
traindset <- traindset[, -c(1:7)]
testdset <- testdset[, -c(1:7)]
traindset <- traindset[, colSums(is.na(traindset)) == 0]
testdset <- testdset[, colSums(is.na(testdset)) == 0]
set.seed(7826) 
partn <- createDataPartition(traindset$classe , p=0.7 ,list=FALSE)
trainpart <- traindset[partn, ]
validpart <- traindset[-partn, ] 
```
### Create models using the training data for determining the accuracy
```{r models, include=TRUE}
set.seed(7826) 
control <- trainControl(method="cv", number=5)
fit_rpart <- train(classe~., data=trainpart, method="rpart", metric ="Accuracy",trControl=control)
fit_svm <- train(classe~., data=trainpart, method="svmRadial", metric ="Accuracy",trControl=control)
```
### Select the best model
```{r best,  fig.width=8,fig.height=6,include=TRUE}
set.seed(7826) 
results <- resamples(list(cart=fit_rpart, svm=fit_svm))
summary(results)
summary(diff(results))
bwplot(results,metric="Accuracy",main="CART Vs SVM")
```

```{r predict, include=TRUE}
##Based on the results SVM is the best model to  be used##
predict_val <- predict(fit_svm, validpart)
confusionMatrix(predict_val, validpart$classe)
```
###Now use SVM model to predict the test data set
```{r predict_test, include=TRUE}
predict_test <- predict(fit_svm, testdset)
print(predict_test)
```
*End of Report*




